A Dependently Typed Multi-Stage Calculus\thanks{Supported by organization x.}

### 2019-05-20 main.tex
Add comments for the outline.
Start wrinting the main technical point.

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bcprules, proof}
\usepackage{fancybox}
\usepackage{mathtools}
\usepackage{float}
\usepackage{xparse}
\usepackage{ebproof}
\usepackage{lscape}
\usepackage{mdframed}

@@

\newcommand{\G}{\Gamma}
\newcommand{\D}{\Delta}
\newcommand{\V}{\vdash}
\newcommand{\VT}{\vdash\hspace{-.50em}\raisebox{0.28em}{\tiny{$\TB$}}}
\newcommand{\iskind}{\text{\ kind}}
\newcommand{\TW}{\triangleright}
\newcommand{\TWL}{\triangleleft}
\newcommand{\F}{\forall}
\newcommand{\TB}{\blacktriangleright}
\newcommand{\TBL}{\blacktriangleleft}
\newcommand{\E}{\equiv}
\newcommand{\FV}{\text{FV}}
\newcommand{\FTV}{\text{FSV}}

@@

\begin{align*}
    \textrm{Type variables} && & X,Y,Z \\
    \textrm{Constants} && & c \\
    \textrm{Variables} && & x,y,z \\
    \textrm{Stage variables} && & \alpha,\beta,\gamma \\
    \textrm{Stage} && & A,B,C \\
    \textrm{Kinds} && K,J,I,H,G & ::= * \mid \Pi x:\tau.K\\
    \textrm{Types} && \tau,\sigma,\rho,\pi,\xi & ::= X \mid \Pi x:\tau.\tau \mid \tau\ M \mid \TW_{\alpha} \tau \mid \F\alpha.\tau \\
    \textrm{Terms} && M,N,L,O,P & ::= c \mid x \mid \lambda x:\tau.M\ \mid M\ M \mid \TB_\alpha M 
    \mid \TBL_\alpha M \mid \Lambda\alpha.M \mid M\ \epsilon \mid \%_\alpha M\\ 
    \textrm{Signature} && \Sigma & ::= \phi \mid X::K \mid c:\tau \\
    \textrm{Contexts} && \Gamma & ::= \phi \mid  \Gamma,x:\tau@A\ (x\not\in\textrm{FV}(\G)) \\
\end{align*}

@@

\begin{align*}
    % \newcommand{\SB}{}
    (\lambda x:\tau.M)[\alpha \mapsto \epsilon] &= \lambda x:\tau[\alpha \mapsto \epsilon].M[\alpha \mapsto \epsilon] \\
    (M\ N)[\alpha \mapsto \epsilon] &= (M[\alpha \mapsto \epsilon])\ (N[\alpha \mapsto \epsilon]) \\
    (\TB_\beta M)[\alpha \mapsto \epsilon] &= \TB_{\beta[\alpha \mapsto \epsilon]} M[\alpha \mapsto \epsilon] \\
    (\TBL_\beta M)[\alpha \mapsto \epsilon] &= \TBL_{\beta[\alpha \mapsto \epsilon]} M[\alpha \mapsto \epsilon] \\
    (\Lambda\beta.M)[\alpha \mapsto \epsilon] &= \Lambda\beta.M[\alpha \mapsto \epsilon] & (\text{if } \alpha \neq \beta) \\
    (\Lambda\beta.M)[\alpha \mapsto \epsilon] &= \Lambda\beta.M & (\text{if } \alpha = \beta) \\
    (M\ \epsilon)[\alpha \mapsto \epsilon] &= M[\alpha \mapsto \epsilon]\ \epsilon \\
    (\%_\beta M)[\alpha \mapsto \epsilon] &= \%_{\beta[\alpha \mapsto \epsilon]}M[\alpha \mapsto \epsilon] 
\end{align*}

@@

% 各論
The rules \TVar, \TAbs, and \TApp are almost same as those in simply typed lambda calculus if you ignore stage annotations.

@@

\begin{definition}[Values]
$A \neq \epsilon$\\
\begin{align*}
    \textrm{Values} && v^\epsilon \in V^\epsilon & ::= \lambda x:\tau.M \mid\ \TB_\alpha v^\alpha \mid \Lambda\alpha.v^\epsilon & \\
                    && v^A \in V^A & ::= x \mid \lambda x:\tau.v^A \mid v^A\ v^A \mid\ \TB_\alpha v^{A\alpha} 
                                           \mid \Lambda\alpha.v^A \mid v^A\ \epsilon &\\
                                    &&& \quad\   \mid\ \TBL_\alpha v^{A'} (\text{if } A'\alpha = A \text{and } A' \neq \epsilon) & \\
                                    &&& \quad\   \mid \%_\alpha v^{A'} (\text{if } A'\alpha = A) & \\
    \textrm{Redexes} && R^\epsilon & ::= (\lambda x:\tau.M)\ v^\epsilon \mid (\Lambda\alpha.v^\epsilon)\ \epsilon & \\
                     && R^\alpha & ::=\ \TBL_\alpha \TB_\alpha M & \\
\end{align*}
\end{definition}

@@

\begin{definition}[Evaluation Context]
$A \neq \epsilon$\\
\begin{align*}
    E^\epsilon_B \in ECtx^\epsilon_B & ::= \square\ (\text{if\ } B = \epsilon) \mid E^\epsilon_B\ M \mid v^e\ E^\epsilon_B
                                           \mid \TB_\alpha E^\alpha_B \mid \Lambda\alpha.E^\epsilon_B
                                           \mid E^\epsilon_B\ \epsilon  \\
    E^A_B \in ECtx^A_B & ::= \square\ (\text{if } A = B) \mid \lambda x:\tau.E^A_B \mid E^A_B\ M \mid v^A\ E^A_B
                                           \mid E^\epsilon_B \mid \TB_\alpha E^{A\alpha}_B
                                           \mid \TBL_\alpha E^{A'}_B \ (\text{where } A'\alpha = A) \\
                                           & \quad \mid \Lambda\alpha.E^\epsilon_B
                                           \mid E^A_B\ \epsilon \mid \%_\alpha\ E^{A'}_B \ (\text{where } A'\alpha = A)\\
\end{align*}
\end{definition}

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bcprules, proof}
\usepackage{fancybox}
\usepackage{mathtools}
\usepackage{float}
\usepackage{xparse}
% \usepackage{ebproof}
\usepackage{lscape}
\usepackage{mdframed}

@@

\newcommand{\G}{\Gamma}
\newcommand{\D}{\Delta}
\newcommand{\V}{\vdash_\Sigma}
\newcommand{\VT}{\vdash\hspace{-.50em}\raisebox{0.28em}{\tiny{$\TB$}}}
\newcommand{\iskind}{\text{\ kind}}
\newcommand{\TW}{\triangleright}
\newcommand{\TWL}{\triangleleft}
\newcommand{\F}{\forall}
\newcommand{\TB}{\blacktriangleright}
\newcommand{\TBL}{\blacktriangleleft}
\newcommand{\E}{\equiv}
\newcommand{\FV}{\text{FV}}
\newcommand{\FTV}{\text{FSV}}

@@

\begin{align*}
    \textrm{Type variables}  &  &                          & X,Y,Z                                                                             \\
    \textrm{Constants}       &  &                          & c                                                                                 \\
    \textrm{Variables}       &  &                          & x,y,z                                                                             \\
    \textrm{Stage variables} &  &                          & \alpha,\beta,\gamma                                                               \\
    \textrm{Stage}           &  &                          & A,B,C                                                                             \\
    \textrm{Kinds}           &  & K,J,I,H,G                & ::= * \mid \Pi x:\tau.K                                                           \\
    \textrm{Types}           &  & \tau,\sigma,\rho,\pi,\xi & ::= X \mid \Pi x:\tau.\tau \mid \tau\ M \mid \TW_{\alpha} \tau \mid \F\alpha.\tau \\
    \textrm{Terms}           &  & M,N,L,O,P                & ::= c \mid x \mid \lambda x:\tau.M\ \mid M\ M \mid \TB_\alpha M 
    \mid \TBL_\alpha M \mid \Lambda\alpha.M \mid M\ \epsilon \mid \%_\alpha M                                                                  \\ 
    \textrm{Signature}       &  & \Sigma                   & ::= \phi \mid X::K \mid c:\tau                                                    \\
    \textrm{Contexts}        &  & \Gamma                   & ::= \phi \mid  \Gamma,x:\tau@A\ (x\not\in\textrm{FV}(\G))                         \\
\end{align*}

@@

\begin{align*}
    % \newcommand{\SB}{}
    (\lambda x:\tau.M)[\alpha \mapsto \epsilon] & = \lambda x:\tau[\alpha \mapsto \epsilon].M[\alpha \mapsto \epsilon]                                  \\
    (M\ N)[\alpha \mapsto \epsilon]             & = (M[\alpha \mapsto \epsilon])\ (N[\alpha \mapsto \epsilon])                                          \\
    (\TB_\beta M)[\alpha \mapsto \epsilon]      & = \TB_{\beta[\alpha \mapsto \epsilon]} M[\alpha \mapsto \epsilon]                                     \\
    (\TBL_\beta M)[\alpha \mapsto \epsilon]     & = \TBL_{\beta[\alpha \mapsto \epsilon]} M[\alpha \mapsto \epsilon]                                    \\
    (\Lambda\beta.M)[\alpha \mapsto \epsilon]   & = \Lambda\beta.M[\alpha \mapsto \epsilon]                            & (\text{if } \alpha \neq \beta) \\
    (\Lambda\beta.M)[\alpha \mapsto \epsilon]   & = \Lambda\beta.M                                                     & (\text{if } \alpha = \beta)    \\
    (M\ \epsilon)[\alpha \mapsto \epsilon]      & = M[\alpha \mapsto \epsilon]\ \epsilon                                                                \\
    (\%_\beta M)[\alpha \mapsto \epsilon]       & = \%_{\beta[\alpha \mapsto \epsilon]}M[\alpha \mapsto \epsilon] 
\end{align*}

@@

% 各論

% 通常のルール
The rules \TVar , \TAbs, and \TApp\ are almost same as those in simply typed lambda calculus if you ignore stage annotations.
The rule \TVar\ means that a variable can appear only at the stage in which it is declared.
The rule \TConst\ means any constants in the signature can appear at any stage.
For example, if we have a signature $\Sigma$ which is $\textrm{bool} :: *, \textrm{true}: \textrm{bool}, \textrm{false}: \textrm{bool}$,
the following derivation tree in Figure \ref{fig:tconst-derivation-tree} is admissible.

\begin{figure}
	\begin{center}
		\begin{minipage}{0.50\hsize}
			\infer[\TConst]
			{\G \V \textrm{true}:\textrm{bool}@\alpha\beta}
			{\textrm{true}:\textrm{bool} \in \Sigma \andalso
				\ID{\G\V\textrm{bool}::*@\alpha\beta} \andalso
			}
			\caption{A derivation tree using \TConst}
			\label{fig:tconst-derivation-tree}
		\end{minipage}
	\end{center}
\end{figure}

% 多段階計算のルール
The rules \TTB, \TTBL, \TGen, \TIns, and \TCsp are rules for a multistage calculus.
They are corresponding to quoting a code, unquoting a code, making a stage abstraction, 
application of $\epsilon$ stage to a stage abstraction, and cross-stage persistence, respectively.
Please check Hanada and Igarashi \cite{Hanada2014} and Tsukada and Igarashi \cite{Tsukada} for details of these rules.

% \TConv
The main technical point of this paper is \TConv and type equivalence rules needed in \TConv.

@@

\begin{figure}
	\begin{center}
		\infrule{\G\V \tau \E \sigma :: *@A \andalso \G,x:\tau@A \V \rho \E \pi :: *@A}{\G\V\Pi x:\tau.\rho \E \Pi x:\sigma.\pi :: *@A}{\QTAbs} \andalso
		\infrule{\G\V \tau \E \sigma :: (\Pi x:\rho.K)@A \andalso \G\V M \E N : \rho @A}{\G\V \tau\ M \E \sigma\ N :: K[x \mapsto M]@A}{\QTApp} \\[2mm]
		\infrule{\G\V \tau \E \sigma :: *@{A\alpha}}{\G\V \TW_{\alpha} \tau \E \TW_{\alpha} \sigma :: *@A}{\textsc{QT-$\TW$}}\andalso
		\infrule{\G\V \tau \E \sigma :: *@A \andalso \alpha\notin\rm{FTV}(\G)\cup\rm{FTV}(A)}{\G\V \forall\alpha.\tau \E  \forall\alpha.\sigma :: *@A}{\textsc{QT-Gen}} \andalso
		\infrule{\G\V \tau \E \sigma :: K@A}{\G\V \tau \E \sigma :: K@{A\alpha}}{\textsc{QT-Csp}} \\[4mm]
		\infrule{\G\V \tau::K@A}{\G\V \tau\E\tau :: K@A}{\textsc{QT-Refl}} \andalso
		\infrule{\G\V \tau \E \sigma :: K@A}{\G\V \sigma \E \tau :: K@A}{\textsc{QT-Sym}} \andalso
		\infrule{\G\V \tau \E \sigma :: K@A \andalso \G\V \sigma \E \rho  :: K@A}{\G\V \tau \E \rho  :: K@A}{\textsc{QT-Trans}} \\[4mm]
		\caption{Type Equivalence Rules}
		\label{fig:type-equivalence-rules}
	\end{center}
\end{figure}

% Normal Conversion Rules

% \QCsp

% Example of QCsp (mulmat)

\begin{definition}[Values]
	$A \neq \epsilon$\\
	\begin{align*}
		\textrm{Values}  &  & v^\epsilon \in V^\epsilon & ::= \lambda x:\tau.M \mid\ \TB_\alpha v^\alpha \mid \Lambda\alpha.v^\epsilon                                       & \\
		                 &  & v^A \in V^A               & ::= x \mid \lambda x:\tau.v^A \mid v^A\ v^A \mid\ \TB_\alpha v^{A\alpha} \mid \Lambda\alpha.v^A \mid v^A\ \epsilon & \\
		                 &  &                           & \quad\   \mid\ \TBL_\alpha v^{A'} (\text{if } A'\alpha = A \text{and } A' \neq \epsilon)                           & \\
		                 &  &                           & \quad\   \mid \%_\alpha v^{A'} (\text{if } A'\alpha = A)                                                           & \\
		\textrm{Redexes} &  & R^\epsilon                & ::= (\lambda x:\tau.M)\ v^\epsilon \mid (\Lambda\alpha.v^\epsilon)\ \epsilon                                       & \\
		                 &  & R^\alpha                  & ::=\ \TBL_\alpha \TB_\alpha M                                                                                      & \\
	\end{align*}

	% Valueの説明

\end{definition}

@@

\begin{definition}[Evaluation Context]
	$A \neq \epsilon$\\
	\begin{align*}
		E^\epsilon_B \in ECtx^\epsilon_B & ::= \square\ (\text{if\ } B = \epsilon) \mid E^\epsilon_B\ M \mid v^e\ E^\epsilon_B
		\mid \TB_\alpha E^\alpha_B \mid \Lambda\alpha.E^\epsilon_B
		\mid E^\epsilon_B\ \epsilon                                                                                                \\
		E^A_B \in ECtx^A_B               & ::= \square\ (\text{if } A = B) \mid \lambda x:\tau.E^A_B \mid E^A_B\ M \mid v^A\ E^A_B
		\mid E^\epsilon_B \mid \TB_\alpha E^{A\alpha}_B
		\mid \TBL_\alpha E^{A'}_B \ (\text{where } A'\alpha = A)                                                                   \\
		                                 & \quad \mid \Lambda\alpha.E^\epsilon_B
		\mid E^A_B\ \epsilon \mid \%_\alpha\ E^{A'}_B \ (\text{where } A'\alpha = A)                                               \\
	\end{align*}
\end{definition}
